Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-03-02 doc/tfp-2011/main.tex
updated (not yet complete) related work subsection [KK]

There are several implementations of Google's Map/Reduce in functional
programming languages. But Functory was just inspired by Google's
Map/Reduce and is not exactly a Map/Reduce implementation. The
simplest difference comes from the very fact that Functory does not
operate on key/value pairs. In order to compare and distinguish
Functory from other work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item Distributed Functional Languages (DFLs) - functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, JoCaml, Glasgow Distributed
  Haskell, Erlang, etc.
\item Libraries for existing and well-known functional programming
  languages that could be used in order to relieve the user from
  implementing intricate details of distribution like
  fault-tolerance, socket programming, etc.
\end{enumerate}

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for transmitting computations. On the other hand,
\JoCaml\ can not in a true sense be used as a distributed computing
library. It does not provide fault tolerance, which is indispensable
in a distributed setting. The user has to include code for fault
tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}.

ML5~\cite{ML5}, a variant of ML, is a programming language for
distributed computing, specialized for web programming. It provides
primitives for transferring control between the client and the server,
as well as low-level primitives for marshalling the data. In
particular, it achieves type-safe marshalling. As in the case before,
ML5 is a programming language that offers primitives for code
mobility, rather than a user-friendly API for distribution of
computation and fault-tolerance.

Glasgow Distributed Haskell

The above mentioned languages are programming languages themselves,
our library, Functory, could be implemented using any of the above.

@@

OcamlP3L

-----------  BELOW THIS IS THE OLD RELATED WORK --------

@@

>>>>

There are several implementations of Google's Map/Reduce in functional
programming languages. But Functory was just inspired by Google's
Map/Reduce and is not exactly a Map/Reduce implementation. The
simplest difference comes from the very fact that Functory does not
operate on key/value pairs. In order to compare and distinguish
Functory from other work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item \textit{Distributed Functional Languages (DFLs)} - functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, JoCaml, Glasgow Distributed
  Haskell, Erlang, etc.
\item \textit{Libraries for existing functional languages} - that
  could be readily used in order to avoid implementing details like
  task distribution, fault-tolerance, socket programming, etc.
\end{enumerate}

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for transmitting computations. On the other hand,
\JoCaml\ can not in a true sense be used as a distributed computing
library. It does not provide fault tolerance, which is indispensable
in a distributed setting. The user has to include code for fault
tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
programming language for distributed computing, specialized for web
programming. It provides primitives for transferring control between
the client and the server, as well as low-level primitives for
marshalling the data. In particular, it achieves type-safe
marshalling. As in the case before, ML5 is a programming language that
offers primitives for code mobility, rather than a user-friendly API
for distribution of computation and fault-tolerance. Glasgow
\textbf{Distributed Haskell: TODO} Erlang~\cite{erlang} is a
programming language which has features for distribution and fault
tolerance. It has features for task distribution and is more
well-known for its rich error detection primitives and the ability to
support hot-swapping. The error detection primitives of Erlang allow
nodes to monitor processes in other nodes and also facilitate
automatic migration of tasks in failed nodes to recovered or active
nodes. CamlP3l~\cite{camlP3l} mixes the features of functional
programming with predefined patterns for parallel computation to offer
a parallel programming environment. Again, it is a programming
language offering primitives for distributing computation to parallel
processes and also to merge the results from parallel executions. The
above mentioned languages are programming languages themselves, our
library, Functory, could be implemented using any of the above. The
main goal of Functory is to provide the users of an existing
functional programming language high-level user-friendly APIs that
hides the messy details of task distribution and fault-tolerance,
enabling a ready-to-use library for efficient utilization of computing
resources.

@@

PlasmaMR~\cite{plasma} is an implementation of Google's MapReduce on a
distributed file system PlasmaFS. It is able to use PlasmaFS to its
advantage - the ability of the file system to handle large files and
query functions of the functions that implement data locality to
optimize network traffic. % However, Plasma
% MapReduce does not support fault tolerance.
The Disco
project~\cite{disco} is an implementation of Google's MapReduce in
Erlang~\cite{erlang}. Our library is not an impementation of Google's
MapReduce. A MapReduce implementation in \Ocaml\ is Yohann Padioleau's
MapReduce implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is
built on top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a
homemade protocol for message passing.  Currently, we have less
flexibility w.r.t. deployment of the user program than OCamlMPI; on
the other hand, we provide a more generic API together with fault
tolerance.

\begin{comment}

@@

\end{comment}
